Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-11-22 PAPER.tex
-hard wrap on text -plots in paper now have updated analytic to account for lower Tsys value

We begin with our data matrix, $\textbf{x}$, which contains our measured visibilities in Jy. It has length ($N_{t}, N_{f}$), where $N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of the Fourier transform of the sky along $2$ spatial dimensions, and since we are interested in $3$-dimensional Fourier-modes we only need to take one Fourier transform of our data along the line-of-sight dimension. We do this when forming the un-normalized power spectrum estimate $\hat{q}_{\alpha}$:

@@

\noindent The matrix $\textbf{Q}$ is an operator that takes our frequency-domain visibilities and Fourier-transforms them into power spectrum space, while also taking into account physical constants such as the Boltzmann constant used to convert Jy to Kelvin, as well as cosmological scalings. The index $\alpha$ denotes a waveband in $k_{\parallel}$, where $k_{\parallel}$ is the Fourier-dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}). $\textbf{R}$ is a weighting matrix --- as an example, inverse covariance weighting would set $\textbf{R} = \textbf{C}^{-1}$ and an unweighted case would use $\textbf{R} = \textbf{I}$, the identity matrix.

@@

\noindent where $\hat{\textbf{p}}$ is the normalized estimate of the true power spectrum. The data analyst has a choice for $\textbf{M}$ --- for simplicity in this section we set $\textbf{M} \propto \textbf{I}$, although we explore other cases for the analysis of PAPER-64 data as explained in Section \ref{sec:Bias}.

@@

We create mock visibilities, $\textbf{x}$, and assume a non-tracking, drift scan observation. Hence, flat spectrum sources (away from zenith) oscillate in time and frequency in our measurements. We therefore form a mock bright foreground signal, $\textbf{x}_{FG}$, as a complex sinusoid that varies smoothly in time and frequency, a simplistic but realistic representation of a single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. A more realistic EoR signal would have a sloped power spectrum in $\hat{\textbf{p}}$(k) (instead of flat, as in the case of white noise), which could be simulated by introducing frequency correlations into the mock EoR signal. However, in this paper we treat all $k$'s separately, so a simplistic white noise approximation can be used. Our combined data vector is then $\textbf{x} = \textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure \ref{fig:toy_sigloss1}.

@@

\noindent assuming $\langle\textbf{x}\rangle = 0$ (a reasonable assumption since fringes average to $0$ over a sufficient amount of time), where $\langle \rangle$ denotes a finite average over time $t$. The inverse covariance matrix is therefore $\textbf{R} = \hat{\textbf{C}}^{-1}$.

First, we compute the power spectrum of $\textbf{x}$ using OQE formalism and a weighting matrix of $\hat{\textbf{C}}^{-1}$. The result is shown in green in the left plot of Figure \ref{fig:toy_sigloss3} (the right plot shows an alternate weighting approach and will be discussed further below). Also plotted in the figure are the unweighted ($\textbf{R} = \textbf{I}$) power spectrum of $\textbf{x}_{FG}$ (blue) and $\textbf{x}_{EoR}$ (red).

@@

If $\textbf{C}$ is computed from the data itself, it carries the risk of over-fitting information in the data and introducing a multiplicative bias (per $k$) to estimates of the signal. For a mathematical derivation of signal loss arising from a data-estimated covariance matrix, see Appendix \ref{sec:sigloss_appendix}. Here we will describe the origin of this signal loss intuitively.

It turns out that because we estimated $\hat{\textbf{C}}$ from our data, its eigenspectrum differs from the eigenspectrum of our true $\textbf{C}$, and this difference has consequences on our result. An eigenspectrum ranks the eigenvalues of a matrix from highest to lowest and can be thought of as a spectrum of weights that are given to each spectral mode in the data. In other words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. This is usually not the case for a covariance matrix, for which a sloped eigenspectrum means that modes are given different weights. The modes with the highest eigenvalues are down-weighted the most.

@@

In general, the strongest modes of $\hat{\textbf{C}}$ (highest eigenvalues) are expected to be dominated by bright foregrounds --- the most prominent shapes in the data. For these `strong' modes, where foregrounds outshine the EoR signal, $\hat{\textbf{C}} \sim \textbf{C}$ and down-weighting these modes is beneficial. This is demonstrated in the toy model by the successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid shape and assigns it the highest eigenvalue (the peak in Figure \ref{fig:toy_sigloss2}). In Figure \ref{fig:toy_sigloss12} we show the covariances of our toy model datasets along with inverse covariance weighted data. The foreground sinusoid is clearly visible in $\hat{\textbf{C}}_{FG}$ but effectively absent from $\hat{\textbf{C}}^{-1}_{FG}\textbf{x}_{FG}$.

@@

The danger of an empirically estimated covariance matrix comes from not being able to describe the `weak' eigenmodes of $\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. If these weak eigenmodes of $\hat{\textbf{C}}$ characterize EoR modes, the EoR signal will be down-weighted (not as much as the strong foreground modes, but still down-weighted). This leads to the `overfitting' of noise (or `overfitting' of signal), which is signal loss.

@@

As a first test, we model the covariance matrix of EoR as a proof of concept that if perfect models are known, signal loss can be avoided. We know that our simulated EoR signal should have a covariance matrix that mimics the identity matrix, with its variance encoded along the diagonal. We model $\textbf{C}_{EoR}$ as such, instead of computing it based on $\textbf{x}_{EoR}$ itself. Next, we add $\textbf{C}_{EoR} + \hat{\textbf{C}}_{FG}$ (where $\hat{\textbf{C}}_{FG} = \langle\textbf{x}_{FG}\textbf{x}_{FG}^{\dagger}\rangle$) to obtain a final $\hat{\textbf{C}}$ to use in weighting. In Figure \ref{fig:toy_sigloss8} (upper left), we see that there is negligible signal loss. This is because by modeling $\textbf{C}_{EoR}$, we avoid over-fitting fluctuations in the data that our model doesn't know about (but an empirically derived $\hat{\textbf{C}}$ would). This is evident when comparing the green and red curves in Figure \ref{fig:toy_sigloss14}. We don't see perfect EoR recovery though, as the green and red curves vary slightly in Figure \ref{fig:toy_sigloss8}. This deviation is a consequence of the difference between the true and modeled covariance matrices. In practice such a weighting option is not feasible, as we do not know $\textbf{C}_{EoR}$.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss14.png}
	\caption{We compare the eigenspectrum of an empirically calculated $\hat{\textbf{C}}$ (green) to that of four alternate weighting options, including modeling the covariance matrix of EoR (red), regularizing $\hat{\textbf{C}}$ by adding an identity matrix to it (blue), using only the first three eigenmodes of $\hat{\textbf{C}}$ (yellow), and multiplying an identity matrix with $\hat{\textbf{C}}$ (magenta). All eigenspectra (except the green) are relatively flat and don't exhibit signal loss. All were computed for the fringe-rate filtered toy model case.}
	\label{fig:toy_sigloss14}
\end{figure}

@@

The last regularization scheme we are highlighting here is setting $\hat{\textbf{C}} = \hat{\textbf{C}} \circ \textbf{I}$ (element-wise multiplication), or inverse variance weighting (keeping only the diagonal elements of $\hat{\textbf{C}}$). In the bottom right panel of Figure \ref{fig:toy_sigloss8}, we see that this method does a poor job down-weighting foregrounds. For this toy model, our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by the identity, element-wise, results in a diagonal matrix, meaning we are only left with correlation information between the same two frequencies. Because we disregard information from all other frequency combinations in this case, we do a poor job suppressing the foreground. But because we flattened the eigenspectrum, we also avoid signal loss.

Although the fourth method did not successfully recover EoR for this particular simulation, it is important that we show that there are many options for estimating a covariance matrix, and some may be more effective than others based on the spectral nature of the components in a dataset. One may imagine a situation where a particular systematic is contained to an isolated frequency (such as radio frequency interference or crosstalk). In such a case, preserving only the diagonal elements of $\hat{\textbf{C}}$ would be an effective way of removing this contamination.

@@

This example highlights the importance of understanding how analysis techniques (e.g. fringe-rate filtering) can affect a common statistical procedure like bootstrapping. Bootstrapping as a means of estimating power spectrum errors from real fringe-rate filtered data requires knowledge of the number of independent samples, which is not always a trivial task. For example, computing the effective number of independent samples of fringe-rate filtered data is not as simple as counting the number of averages performed. Down-sampling a time-averaged signal is straightforward using a boxcar average, but non-trivial with a more complicated convolution function that has long tails. Hence, we do not recommend bootstrapping unless the number of independent samples along the axis that is being re-sampled is well-determined.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error1.png}
	\caption{Error estimation produced by bootstrapping as a function of the number of elements drawn per bootstrap when sampling with replacement. The star represents the standard deviation of $500$ bootstraps, each created by drawing $1000$ elements (with replacement) from a length $1000$ array of a Gaussian random signal. The black points correspond to time-averaged data (correlated data) which has $100$ independent samples. They illustrate how errors can be underestimated if drawing more elements than there are independent samples in the data. The estimated errors match up with the theoretical prediction only at $N=100$.}
	\label{fig:toy_error1}
\end{figure}

@@

Because foreground spectra are smooth in frequency, they preferentially show up at low delay, or $k$ modes. For a particular baseline length, there is a maximum delay imposed on foregrounds, which corresponds to the light-crossing time between two antennas in a baseline. For longer baselines, this value increases, producing what is known as ``the wedge" (\citealt{parsons_et_al2012b}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{vedantham_et_al2012}; \citealt{thyagarajan_et_al2013}; \citealt{pober_et_al2013}; \citealt{datta_et_al2010}). The wedge describes a region in $k$-space contaminated by foregrounds, bounded by baseline length (which is proportional to $k_{\perp}$) and delay (which is proportional to $k_{\parallel}$). Properties of the wedge can be used to isolate and remove foregrounds, as done by P64, \citet{parsons_et_al2014}, and \citet{jacobs_et_al2015}.

@@

To mimic the presence of a systematic in part of the measurement, we add a toy sinusoid foreground, similar to the one used in Section \ref{sec:toymodel}, to the first $100$ time integrations of both measurements. This represents a foreground signal present in, for example, the first half of the LST range used for analysis, but not the second half. Mathematically, if  $\textbf{n}$ is noise, $\textbf{e}$ is the EoR signal, and $\textbf{fg}$ is the foreground signal, the two measurements (which are cross-multiplied to form power spectra) can be written as:

@@

We form $3$ different power spectrum estimates shown in Figure \ref{fig:toy_bias1}. The first is a null test where we subtract $\textbf{x}_{2}$ from $\textbf{x}_{1}$ for both measurements $a$ and $b$. This is equivalent to splitting up a full dataset along an axis (in this case, time) and subtracting the two to remove sky signal that should ideally be present in both. We cross-multiply measurements $a$ and $b$ to form an un-biased (thermal noise-wise) estimate (blue curve). The second estimate, shown in red, is the same null test with the foreground systematic removed (eliminate $\textbf{fg}$ in Equations \ref{eq:bias1} and \ref{eq:bias2}). Finally, we also show the noise power spectrum (green).

@@

While the null test is useful for testing noise properties and the uniformity of a dataset, jackknives are useful in pinpointing which data subsets are contaminated by biases and which are not; in our toy model we see that the bias exists only in $\textbf{x}_{1}$ (Figure \ref{fig:toy_bias2}). If foreground or noise biases exist in a dataset, jackknives can tease them out and provide insight into possible sources. For example, if jackknives along the time-axis reveal a bias present at a certain LST, a likely explanation would be excess foreground emission from a radio source in the sky at that time. A jackknife test involving data before and after the application of a fringe-rate filter can reveal whether cross-talk noise bias is successfully suppressed with the filter, or if similar-shaped detections in both power spectra suggest otherwise. There are many other jackknife axes of which we will not go into detail here, including baseline, frequency, and polarization. Ultimately, an EoR detection should persist through them all and a clean measurement should exhibit noise-like null spectra.

@@

Suppose that $\textbf{e}$ is the injected EoR (at some amplitude level), and $\textbf{x}$ is our data vector. We define $\textbf{r}$ to be the data plus the EoR signal:

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{out,\alpha} &=& \hat{\textbf{p}}_{r,\alpha}-\hat{\textbf{p}}_{x,\alpha} \nonumber \\
&\propto& \textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

@@

\begin{eqnarray}
\label{eq:crossterm}
P_{out,\alpha} &\propto& (\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&\propto& \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

@@

However, deeper investigations of these terms reveal that they contain non-negligible power. In fact, foreground modes and signal modes are anti-correlated on average, and this negative cross-term power arises from inverting the covariance matrix $\textbf{C}_{r}$. Mathematically, the covariance of the data plus EoR is:

@@

\noindent where the data-only term (first term) and EoR-only term (last term) are approximately diagonal matrices. The middle two terms, however, have a larger amount of power in their off-diagonal terms relative to their diagonal elements, since $\textbf{e}$ and $\textbf{x}$ are statistically unrelated. Upon inversion, these off-diagonal terms become negative, driving the cross-terms in Equation \ref{eq:pout_expand} to have negative power. \cc{someone help me write this more elegantly...}This effect is explained in more depth in \citet{switzer_et_al2015}.

@@

\begin{eqnarray}
P_{in,\alpha} &\propto& \textbf{e}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e} \\
P_{out,\alpha} &\propto& \textbf{r}^{\dagger}\textbf{C}_{r}^{-1}\textbf{Q}^{\alpha}\textbf{C}_{r}^{-1}\textbf{r} - \textbf{x}^{\dagger}\textbf{C}_{x}^{-1}\textbf{Q}^{\alpha}\textbf{C}_{x}^{-1}\textbf{x} 
\end{eqnarray}

@@

Finally, the solid grey horizontal lines in Figure \ref{fig:sigloss_transfercurve} mark the peak of the data distribution $\hat{\textbf{p}}$ (the distribution is created from the $20$ bootstraps) for both the inverse covariance weighted and unweighted cases. One can eyeball that a data value of $10^{5} \, mK^{2} (h^{-1} Mpc)^{3}$, for example, would map approximately to a value of $\sim10^{8} \, mK^{2} (h^{-1} Mpc)^{3}$ using the transfer curve for the inverse covariance weighted case, implying a signal loss factor of $\sim1000$.

@@

One peculiar aspect of Figure \ref{fig:sigloss_transfercurve} is the fact that at low $P_{in}$ values it appears that we can have signal gain ($P_{out} > P_{in}$). This is unphysical in nature but caused due to the dominating cross-terms involving $\textbf{e}$ and $\textbf{x}$ in Equation \eqref{eq:pout_expand} once $\textbf{e}$ becomes small. As $\textbf{e}$ increases, we move into a regime where $P_{out} \sim P_{in}$, and then eventually into a regime where $P_{out} < P_{in}$ when $\textbf{e}$ is large enough to be destroyed if weighting the data using itself. Although we only show figures for one $k$ value, we note that the shape of the transfer curve is nearly identical for all $k$'s (though we treat each $k$ separately).

@@

Our revised sensitivity estimate for PAPER-64 is shown in comparison with that of P64 in Figure \ref{fig:sense_check}. Together, the revised parameters yield a decrease in sensitivity (higher noise floor) by a factor of $\sim7$.

>>>>

We begin with our data matrix, $\textbf{x}$, which contains our measured visibilities in Jy. It has length ($N_{t}, N_{f}$), where 
$N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of 
the Fourier transform of the sky along $2$ spatial dimensions, and since we are interested in $3$-dimensional Fourier-modes 
we only need to take one Fourier transform of our data along the line-of-sight dimension. We do this when forming the un-
normalized power spectrum estimate $\hat{q}_{\alpha}$:

@@

\noindent The matrix $\textbf{Q}$ is an operator that takes our frequency-domain visibilities and Fourier-transforms them into 
power spectrum space, while also taking into account physical constants such as the Boltzmann constant used to convert Jy to 
Kelvin, as well as cosmological scalings. The index $\alpha$ denotes a waveband in $k_{\parallel}$, where $k_{\parallel}$ is the 
Fourier-dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}). $\textbf{R}$ is a weighting matrix --- as 
an example, inverse covariance weighting would set $\textbf{R} = \textbf{C}^{-1}$ and an unweighted case would use $
\textbf{R} = \textbf{I}$, the identity matrix.

@@

\noindent where $\hat{\textbf{p}}$ is the normalized estimate of the true power spectrum. The data analyst has a choice for $
\textbf{M}$ --- for simplicity in this section we set $\textbf{M} \propto \textbf{I}$, although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}.

@@

We create mock visibilities, $\textbf{x}$, and assume a non-tracking, drift scan observation. Hence, flat spectrum sources (away 
from zenith) oscillate in time and frequency in our measurements. We therefore form a mock bright foreground signal, $
\textbf{x}_{FG}$, as a complex sinusoid that varies smoothly in time and frequency, a simplistic but realistic representation of a 
single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. A more 
realistic EoR signal would have a sloped power spectrum in $\hat{\textbf{p}}$(k) (instead of flat, as in the case of white noise), 
which could be simulated by introducing frequency correlations into the mock EoR signal. However, in this paper we treat all 
$k$'s separately, so a simplistic white noise approximation can be used. Our combined data vector is then $\textbf{x} = 
\textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure 
\ref{fig:toy_sigloss1}.

@@

\noindent assuming $\langle\textbf{x}\rangle = 0$ (a reasonable assumption since fringes average to $0$ over a sufficient 
amount of time), where $\langle \rangle$ denotes a finite average over time $t$. The inverse covariance matrix is therefore $
\textbf{R} = \hat{\textbf{C}}^{-1}$.

First, we compute the power spectrum of $\textbf{x}$ using OQE formalism and a weighting matrix of $\hat{\textbf{C}}^{-1}$. 
The result is shown in green in the left plot of Figure \ref{fig:toy_sigloss3} (the right plot shows an alternate weighting approach 
and will be discussed further below). Also plotted in the figure are the unweighted ($\textbf{R} = \textbf{I}$) power spectrum of $
\textbf{x}_{FG}$ (blue) and $\textbf{x}_{EoR}$ (red).

@@

If $\textbf{C}$ is computed from the data itself, it carries the risk of over-fitting information in the data and introducing a 
multiplicative bias (per $k$) to estimates of the signal. For a mathematical derivation of signal loss arising from a data-
estimated covariance matrix, see Appendix \ref{sec:sigloss_appendix}. Here we will describe the origin of this signal loss 
intuitively.

It turns out that because we estimated $\hat{\textbf{C}}$ from our data, its eigenspectrum differs from the eigenspectrum of our 
true $\textbf{C}$, and this difference has consequences on our result. An eigenspectrum ranks the eigenvalues of a matrix from 
highest to lowest and can be thought of as a spectrum of weights that are given to each spectral mode in the data. In other 
words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $
\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. This is usually not the case for a covariance matrix, 
for which a sloped eigenspectrum means that modes are given different weights. The modes with the highest eigenvalues are 
down-weighted the most.

@@

In general, the strongest modes of $\hat{\textbf{C}}$ (highest eigenvalues) are expected to be dominated by bright foregrounds 
--- the most prominent shapes in the data. For these `strong' modes, where foregrounds outshine the EoR signal, $
\hat{\textbf{C}} \sim \textbf{C}$ and down-weighting these modes is beneficial. This is demonstrated in the toy model by the 
successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid shape and 
assigns it the highest eigenvalue (the peak in Figure \ref{fig:toy_sigloss2}). In Figure \ref{fig:toy_sigloss12} we show the 
covariances of our toy model datasets along with inverse covariance weighted data. The foreground sinusoid is clearly visible in 
$\hat{\textbf{C}}_{FG}$ but effectively absent from $\hat{\textbf{C}}^{-1}_{FG}\textbf{x}_{FG}$.

@@

The danger of an empirically estimated covariance matrix comes from not being able to describe the `weak' eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. If these weak eigenmodes of $\hat{\textbf{C}}$ 
characterize EoR modes, the EoR signal will be down-weighted (not as much as the strong foreground modes, but still down-
weighted). This leads to the `overfitting' of noise (or `overfitting' of signal), which is signal loss.

@@

As a first test, we model the covariance matrix of EoR as a proof of concept that if perfect models are known, signal loss can be 
avoided. We know that our simulated EoR signal should have a covariance matrix that mimics the identity matrix, with its 
variance encoded along the diagonal. We model $\textbf{C}_{EoR}$ as such, instead of computing it based on $\textbf{x}
_{EoR}$ itself. Next, we add $\textbf{C}_{EoR} + \hat{\textbf{C}}_{FG}$ (where $\hat{\textbf{C}}_{FG} = \langle\textbf{x}_{FG}
\textbf{x}_{FG}^{\dagger}\rangle$) to obtain a final $\hat{\textbf{C}}$ to use in weighting. In Figure \ref{fig:toy_sigloss8} (upper 
left), we see that there is negligible signal loss. This is because by modeling $\textbf{C}_{EoR}$, we avoid over-fitting 
fluctuations in the data that our model doesn't know about (but an empirically derived $\hat{\textbf{C}}$ would). This is evident 
when comparing the green and red curves in Figure \ref{fig:toy_sigloss14}. We don't see perfect EoR recovery though, as the 
green and red curves vary slightly in Figure \ref{fig:toy_sigloss8}. This deviation is a consequence of the difference between the 
true and modeled covariance matrices. In practice such a weighting option is not feasible, as we do not know $\textbf{C}_{EoR}
$.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss14.png}
	\caption{We compare the eigenspectrum of an empirically calculated $\hat{\textbf{C}}$ (green) to that of four alternate 
weighting options, including modeling the covariance matrix of EoR (red), regularizing $\hat{\textbf{C}}$ by adding an identity 
matrix to it (blue), using only the first three eigenmodes of $\hat{\textbf{C}}$ (yellow), and multiplying an identity matrix with $
\hat{\textbf{C}}$ (magenta). All eigenspectra (except the green) are relatively flat and don't exhibit signal loss. All were 
computed for the fringe-rate filtered toy model case.}
	\label{fig:toy_sigloss14}
\end{figure}

@@

The last regularization scheme we are highlighting here is setting $\hat{\textbf{C}} = \hat{\textbf{C}} \circ \textbf{I}$ (element-
wise multiplication), or inverse variance weighting (keeping only the diagonal elements of $\hat{\textbf{C}}$). In the bottom right 
panel of Figure \ref{fig:toy_sigloss8}, we see that this method does a poor job down-weighting foregrounds. For this toy model, 
our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by 
the identity, element-wise, results in a diagonal matrix, meaning we are only left with correlation information between the same 
two frequencies. Because we disregard information from all other frequency combinations in this case, we do a poor job 
suppressing the foreground. But because we flattened the eigenspectrum, we also avoid signal loss.

Although the fourth method did not successfully recover EoR for this particular simulation, it is important that we show that there 
are many options for estimating a covariance matrix, and some may be more effective than others based on the spectral nature 
of the components in a dataset. One may imagine a situation where a particular systematic is contained to an isolated 
frequency (such as radio frequency interference or crosstalk). In such a case, preserving only the diagonal elements of $
\hat{\textbf{C}}$ would be an effective way of removing this contamination.

@@

This example highlights the importance of understanding how analysis techniques (e.g. fringe-rate filtering) can affect a 
common statistical procedure like bootstrapping. Bootstrapping as a means of estimating power spectrum errors from real 
fringe-rate filtered data requires knowledge of the number of independent samples, which is not always a trivial task. For 
example, computing the effective number of independent samples of fringe-rate filtered data is not as simple as counting the 
number of averages performed. Down-sampling a time-averaged signal is straightforward using a boxcar average, but non-
trivial with a more complicated convolution function that has long tails. Hence, we do not recommend bootstrapping unless the 
number of independent samples along the axis that is being re-sampled is well-determined.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error1.png}
	\caption{Error estimation produced by bootstrapping as a function of the number of elements drawn per bootstrap when 
sampling with replacement. The star represents the standard deviation of $500$ bootstraps, each created by drawing $1000$ 
elements (with replacement) from a length $1000$ array of a Gaussian random signal. The black points correspond to time-
averaged data (correlated data) which has $100$ independent samples. They illustrate how errors can be underestimated if 
drawing more elements than there are independent samples in the data. The estimated errors match up with the theoretical 
prediction only at $N=100$.}
	\label{fig:toy_error1}
\end{figure}

@@

Because foreground spectra are smooth in frequency, they preferentially show up at low delay, or $k$ modes. For a particular 
baseline length, there is a maximum delay imposed on foregrounds, which corresponds to the light-crossing time between two 
antennas in a baseline. For longer baselines, this value increases, producing what is known as ``the 
wedge" (\citealt{parsons_et_al2012b}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{vedantham_et_al2012}; 
\citealt{thyagarajan_et_al2013}; \citealt{pober_et_al2013}; \citealt{datta_et_al2010}). The wedge describes a region in $k$-
space contaminated by foregrounds, bounded by baseline length (which is proportional to $k_{\perp}$) and delay (which is 
proportional to $k_{\parallel}$). Properties of the wedge can be used to isolate and remove foregrounds, as done by P64, 
\citet{parsons_et_al2014}, and \citet{jacobs_et_al2015}.

@@

To mimic the presence of a systematic in part of the measurement, we add a toy sinusoid foreground, similar to the one used in 
Section \ref{sec:toymodel}, to the first $100$ time integrations of both measurements. This represents a foreground signal 
present in, for example, the first half of the LST range used for analysis, but not the second half. Mathematically, if  $\textbf{n}$ 
is noise, $\textbf{e}$ is the EoR signal, and $\textbf{fg}$ is the foreground signal, the two measurements (which are cross-
multiplied to form power spectra) can be written as:

@@

We form $3$ different power spectrum estimates shown in Figure \ref{fig:toy_bias1}. The first is a null test where we subtract $
\textbf{x}_{2}$ from $\textbf{x}_{1}$ for both measurements $a$ and $b$. This is equivalent to splitting up a full dataset along 
an axis (in this case, time) and subtracting the two to remove sky signal that should ideally be present in both. We cross-
multiply measurements $a$ and $b$ to form an un-biased (thermal noise-wise) estimate (blue curve). The second estimate, 
shown in red, is the same null test with the foreground systematic removed (eliminate $\textbf{fg}$ in Equations \ref{eq:bias1} 
and \ref{eq:bias2}). Finally, we also show the noise power spectrum (green).

@@

While the null test is useful for testing noise properties and the uniformity of a dataset, jackknives are useful in pinpointing 
which data subsets are contaminated by biases and which are not; in our toy model we see that the bias exists only in $
\textbf{x}_{1}$ (Figure \ref{fig:toy_bias2}). If foreground or noise biases exist in a dataset, jackknives can tease them out and 
provide insight into possible sources. For example, if jackknives along the time-axis reveal a bias present at a certain LST, a 
likely explanation would be excess foreground emission from a radio source in the sky at that time. A jackknife test involving 
data before and after the application of a fringe-rate filter can reveal whether cross-talk noise bias is successfully suppressed 
with the filter, or if similar-shaped detections in both power spectra suggest otherwise. There are many other jackknife axes of 
which we will not go into detail here, including baseline, frequency, and polarization. Ultimately, an EoR detection should persist 
through them all and a clean measurement should exhibit noise-like null spectra.

@@

Suppose that $\textbf{e}$ is the injected EoR (at some amplitude level), and $\textbf{x}$ is our data vector. We define $\textbf{r}
$ to be the data plus the EoR signal:

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{out,\alpha} &=& \hat{\textbf{p}}_{r,\alpha}-\hat{\textbf{p}}_{x,\alpha} \nonumber \\
&\propto& \textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}
^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

@@

\begin{eqnarray}
\label{eq:crossterm}
P_{out,\alpha} &\propto& (\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) - 
\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&\propto& \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} + \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}
^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

@@

However, deeper investigations of these terms reveal that they contain non-negligible power. In fact, foreground modes and 
signal modes are anti-correlated on average, and this negative cross-term power arises from inverting the covariance matrix $
\textbf{C}_{r}$. Mathematically, the covariance of the data plus EoR is:

@@

\noindent where the data-only term (first term) and EoR-only term (last term) are approximately diagonal matrices. The middle 
two terms, however, have a larger amount of power in their off-diagonal terms relative to their diagonal elements, since $
\textbf{e}$ and $\textbf{x}$ are statistically unrelated. Upon inversion, these off-diagonal terms become negative, driving the 
cross-terms in Equation \ref{eq:pout_expand} to have negative power. \cc{someone help me write this more elegantly...}This 
effect is explained in more depth in \citet{switzer_et_al2015}.

@@

\begin{eqnarray}
P_{in,\alpha} &\propto& \textbf{e}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e} \\
P_{out,\alpha} &\propto& \textbf{r}^{\dagger}\textbf{C}_{r}^{-1}\textbf{Q}^{\alpha}\textbf{C}_{r}^{-1}\textbf{r} - \textbf{x}^{\dagger}
\textbf{C}_{x}^{-1}\textbf{Q}^{\alpha}\textbf{C}_{x}^{-1}\textbf{x} 
\end{eqnarray}

@@

Finally, the solid grey horizontal lines in Figure \ref{fig:sigloss_transfercurve} mark the peak of the data distribution $
\hat{\textbf{p}}$ (the distribution is created from the $20$ bootstraps) for both the inverse covariance weighted and unweighted 
cases. One can eyeball that a data value of $10^{5} \, mK^{2} (h^{-1} Mpc)^{3}$, for example, would map approximately to a 
value of $\sim10^{8} \, mK^{2} (h^{-1} Mpc)^{3}$ using the transfer curve for the inverse covariance weighted case, implying a 
signal loss factor of $\sim1000$.

@@

One peculiar aspect of Figure \ref{fig:sigloss_transfercurve} is the fact that at low $P_{in}$ values it appears that we can have 
signal gain ($P_{out} > P_{in}$). This is unphysical in nature but caused due to the dominating cross-terms involving $\textbf{e}
$ and $\textbf{x}$ in Equation \eqref{eq:pout_expand} once $\textbf{e}$ becomes small. As $\textbf{e}$ increases, we move 
into a regime where $P_{out} \sim P_{in}$, and then eventually into a regime where $P_{out} < P_{in}$ when $\textbf{e}$ is 
large enough to be destroyed if weighting the data using itself. Although we only show figures for one $k$ value, we note that 
the shape of the transfer curve is nearly identical for all $k$'s (though we treat each $k$ separately).

@@

Our revised sensitivity estimate for PAPER-64 is shown in comparison with that of P64 in Figure \ref{fig:sense_check}. 
Together, the revised parameters yield a decrease in sensitivity (higher noise floor) by a factor of $\sim5$ in mK$^{2}$.
